Maintenance planning on French military aircraft operations

### 2018-09-17 R/presentations/CLAIO_presentation/presentation.Rmd
updated CLAIO presentation, graphs, simulated data in R and shiny app.

---
title: "Maintenance planning on French military aircraft operations"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
institute: "ISAE SUPAERO"
date: "September 24, 2018"
output: 
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    # slide_level: 2
    # toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    includes:
        in_header: header.tex
bibliography: MFMP_NRP.bib
editor_options: 
  chunk_output_type: console
---

@@

* FMP: Flight and Maintenance Planning problem.
* A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each task requires a certain number $r_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $h_j$ in each period.
* Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H$ units.

@@

* In @Cho2011, US Army aircraft were assigned daily operations over a year to aircraft in order to minimize the maximum number of maintenances.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In @Chastellux2017, an initial version of the model was done, assigning missions instead of only flight hours.
* In @Seif2018, a generalization was done in order to deal with different types of maintenances and an heterogeneus fleet.

@@

## Complexity analysis (1)

Similar to the 'Fixed interval scheduling' problem presented in @Smet2015.

Let $P = {1, ..., n}$ be the set of tasks to be assigned and $E = {1, ..., m}$ the set of employees. Each task $p \in P$ has a duration $u_p$ , a start time $s_p$ and a finish time $f_p$ = $s_p + u_p$ . Each employee $e$ has a set of tasks $P_e \in P$ that he/she can perform. Similarly, for each task $p$, a set $E_p \in E$ exists, which contains all employees that can perform task $p$. Both $T_e$ and $E_p$ are defined based on qualifications, time windows of tasks and availabilities of employees.

## Complexity analysis (2)

```{r, echo=FALSE, out.width='50%'}

if (show_gantt){
    htmltools::div(class = "row",
        htmltools::div(task_gantt, class = "col-md-6")
        ,htmltools::div(task_employees, class = "col-md-6")
        )
} else {
    c('../img/tasks-days.png', '../img/task-employees.png') %>% paste0(path_to_R_dir, .) %>% 
    knitr::include_graphics
    # knitr::include_graphics('../img/task-employees.png' %>% paste0(path_to_R_dir, .))
}

@@

## Model: objective function

@@

\begin{align}
    & \text{Min}\; W_1 m_{max} + W_2 u_{max}
\end{align}

@@

The two components of the objective function: the max number of maintenance and the max number of unavailable aircraft.
\begin{align}
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq m_{max}
    &t \in \mathcal{T} \\
    &\sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t + D_t\leq u_{max}
    &t \in \mathcal{T}
\end{align}

@@

Mission's needs and incompatibility of several tasks or maintenance in the same period.
\begin{align}
&\sum_{i \in \mathcal{I}_j} a_{jti} = R_j
&j \in \mathcal{J}, t \in \mathcal{T}_j\\
&\sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1
& t \in \mathcal{T}, i \in \mathcal{I}
\end{align}

<!-- ## Model: flow constraints -->

<!-- Visualization of accumulated time constraints for one resource. -->

<!-- ```{r out.width="250px", echo=FALSE} -->
<!-- # img <- readPNG("./../../img/balance2.png") -->
<!-- #  grid.raster(img) -->
<!-- knitr::include_graphics("./../../img/balance2.png") -->
<!-- ``` -->

@@

Time usage of an aircraft depending on the assignment to a task.
<!-- <div class="incremental" style="margin-left: 4em; position: relative">  -->
<!--   <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->
\begin{align}
& rut_{it} \leq rut_{it-1} + H m_{it} - \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j & t =1, ..., \mathcal{T}, i \in \mathcal{I}\\
& rut_{i0} = Rut^{Init}_i
        & i \in \mathcal{I}\\
& rut_{it} \geq H m_{it}
        & t \in \mathcal{T}, i \in \mathcal{I}\\
& rut_{it} \in [0,H]
        & t \in \mathcal{T}, i \in \mathcal{I} \\
& \sum_{i \in \mathcal{I}} rut_{it} \geq Rut^{Init}_{sum}
        & t = |\mathcal{T}|
\end{align}

  <!-- </div>  -->
  <!-- <div style="position: absolute; left: 0; top: 0;background-color:white;width:100%"> -->

# Results

## Experiments

## Sample solution progress

# Current and next steps

## Next steps

* **Add more constraints**: storage of aircraft.
* **Add robustness**: duration of maintenances, new /different missions, unavailable aircraft.

@@

<!-- \bibliographystyle{dcu} -->
<!-- \bibliography{./../biblio/MFMP} -->

>>>>

---
title: "Maintenance planning on French military aircraft operations"
author: |
    | Franco Peschiera, Alain Haït, Olga Battaïa, Nicolas Dupin
institute: "ISAE SUPAERO"
date: "September 24, 2018"
output: 
  slidy_presentation:
    css: ./../../../css/baobab_slidy.css
    incremental: yes
    footer: "ISAE-SUPAERO, Université de Toulouse, France"
    # slide_level: 2
    # toc: true
  beamer_presentation:
    theme: "Berlin"
    toc: true
    slide_level: 2
    incremental: yes
    fig_height: 4
    # citation_package: natbib
    # footer: "ISAE-SUPAERO, Université de Toulouse, France"
    includes:
        in_header: header.tex
bibliography: MFMP_NRP.bib
# nocite: '@*'
editor_options: 
  chunk_output_type: console
---

@@

* FMP: Flight and Maintenance Planning problem.
* A series of $j \in \mathcal{J}$ tasks are planned along a horizon divided into $t \in \mathcal{T}$ periods. Since all tasks are already scheduled, we know time periods $T_j \subset \mathcal{T}$ in which they will be realized. 
* Each task requires a certain number $R_{j}$ of resources $i \in \mathcal{I}$ which it employs for a time duration defined by $H_j$ in each period and needs to be assigned for a minimum of $MT_j$ consecutive periods if assigned.
* Resources require recurrent preventive maintenance operations since the realization of tasks diminish their remaining usage time.
* A maintenance operation takes exactly $m$ periods and cannot be interrupted. It restores the resource's remaining usage time to exactly $H^M$ units.

@@

* In @Cho2011, US Army aircraft were assigned daily operations over a year to minimize the number of maintenances.
* In @Kozanidis2008, Greek aircraft had monthly assignments of maintenances and flight hours in order to maximize the availability and final state of squadrons.
* In @Verhoeff2015, monthly assignments were done and several objectives were taken into account: availability, serviceability and final state.
* In @Seif2018, a generalization of @Kozanidis2008 was done in order to deal with different types of maintenances and an heterogeneus fleet.

@@

# Complexity analysis

## Complexity analysis

Reduction to the 'Fixed interval scheduling' in @Smet2015.

* $P$ tasks with duration $u_p$ and start time $s_p$.
* $E$ employees with a set of tasks $P_e \in P$ that each can perform.
* Minimize the number of required employees to do all tasks.

```{r, echo=FALSE, out.width='45%'}

if (show_gantt){
    htmltools::div(class = "row",
        htmltools::div(task_employees, class = "col-md-6")
        ,htmltools::div(task_gantt, class = "col-md-6")
        )
} else {
    c('../img/tasks-days.png', '../img/task-employees.png') %>% 
        paste0(path_to_R_dir, .) %>%
        knitr::include_graphics()
    # knitr::include_graphics('../img/task-employees.png' %>% paste0(path_to_R_dir, .))
}

@@

@@

\begin{align}
& \text{Min}\; \sum_{t \in \mathcal{T}, i \in \mathcal{I}} m_{it} \times H^{M} - \sum_{i \in \mathcal{I}} rut_{i|\mathcal{T}|}
\end{align}

The objective is to minimize the total number of maintenances and, at the same time, maximize the end status of the resources.

@@

\begin{align}
    & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}} m_{it'} + N_t \leq C^{max}
      & t \in \mathcal{T} \label{eq:capacity1}\\
    & \sum_{i \in \mathcal{I}_j} a_{jti} \geq R_j
            & j \in \mathcal{J}, t \in \mathcal{T}_j  \label{eq:taskres}\\
    & \sum_{t' \in \mathcal{T}^{s}_t} m_{it'} + \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} \leq 1 
            & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:state}
\end{align}

Mission's needs and incompatibility of several tasks or maintenance in the same period.

@@

## Model: minimum assignment times

<!-- <div class="incremental"> -->

\begin{align}
    & a^s_{jti} \geq a_{jti} - a_{j(t-1)i}
            & t =1, ..., \mathcal{T}, j \in \mathcal{J}_t, i \in \mathcal{I}_j \label{eq:start1} \\
    & a^s_{j0i} \geq a_{j0i} - 1 \!1_{i \in \mathcal{A}^{init}_{j}}
            & j \in \mathcal{J}_0, i \in \mathcal{I}_j \label{eq:start2} \\        
    & \sum_{t' \in \mathcal{T}^{MT}_{jt}} a^s_{jt'i} \leq a_{jti} 
    & j \in \mathcal{J}, t \in \mathcal{T}_j, i \in \mathcal{I}_j \label{eq:start3}
\end{align}

Each task needs a minimum continuous duration at each assignment.

<!-- \pause -->

## Model: maintenances

\begin{align}
    & m_{it'} + m_{it} \leq 1
      & t \in \mathcal{T}, t' \in \mathcal{T}^{m}_t, i \in \mathcal{I}\label{eq:ret_min}\\ 
    & \sum_{t' \in \mathcal{T}^{M}_t} m_{it'} \geq  m_{it}
      & t \in \mathcal{T}, i \in \mathcal{I}\label{eq:ret_max}\\
    & m_{it} = 0
      & t \in \mathcal{T}^{m_{ini}}_i, i \in \mathcal{I}\label{eq:ret_min_init} \\
    & \sum_{t \in \mathcal{T}^{M_{ini}}_i} m_{it} \geq  1 
      & i \in \mathcal{I}\label{eq:ret_max_init}
\end{align}

Maintenances cannot executed too soon or too late with respect to the previous one.

<!-- </div> -->

@@

<!-- Time usage of an aircraft depending on the assignment to a task and remaining usage time per aircraft and period. -->

\begin{align}
 & u_{it} \geq \sum_{j \in \mathcal{J}_t \cap \mathcal{O}_i} a_{jti} H_j 
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage1}\\
 & u_{it} \geq u^{min} (1 - \sum_{t' \in \mathcal{T}^{s}_t} m_{it'})
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage2}\\
 & u_{it} \in [0, \max_j{\{H_j\}}]
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:usage_lower}\\
 & rut_{it} \leq rut_{i(t-1)} + H^{M} m_{it} - u_{it}
    & t =1, ..., \mathcal{T}, i \in \mathcal{I} \label{eq:rut_upper}\\
 & rut_{i0} = Rut^{Init}_i
       & i \in \mathcal{I} \label{eq:rut_initial}\\
 & rut_{it} \geq H^{M} m_{it'}
        & t \in \mathcal{T}, t' \in \mathcal{T}^{s}_t, i \in \mathcal{I}\label{eq:rut_lower}\\
 & rut_{it} \in [0,H^{M}]
        & t \in \mathcal{T}, i \in \mathcal{I} \label{eq:mu}
\end{align}

## Model: availability

\begin{align}
   & \sum_{t' \in \mathcal{T}^{s}_t} \sum_{i \in \mathcal{I}_k} m_{it'} + N^{clust}_{kt} \leq A^{clust}_{kt}
    &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-cluster} \\
   & \sum_{i \in \mathcal{I}_k} rut_{it} \geq H^{clust}_{kt}
    &k \in \mathcal{K}, t \in \mathcal{T} \label{eq:avalaibility-hours-cluster}
\end{align}

Each cluster of aircraft needs a minimum level of availability in number and usage hours.

<!-- # Results -->

<!-- ## Experiments -->

<!-- Instances have been created with the following characteristics: -->

<!-- <div class="incremental"> -->

<!-- * 60 - 200 resources. -->
<!-- * 90 periods. -->
<!-- * Tasks: 3 - 9  -->
<!--     * 12 - 36 periods long -->
<!--     * 2 - 5 resources needed -->

<!-- </div> -->

@@

<!-- \bibliographystyle{dcu} -->
<!-- \bibliography{MFMP_NRP.bib} -->
